Extreme rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill

### 2017-09-22 writeup/Supplement.tex
Use Make much more
Final weather types calculated w/ 500 sims

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/PSI_Var_Explained.pdf}
	\caption{
		Variance explained for each EOF loading of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
    Left figure shows the individual variance explained for each EOF and right figure shows cumulative variance explained.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/WT_Classifiability.pdf}
	\caption{
		Classifiability index (see Methods) calculated for several chosen values of $k$, representing the number of weather types used for clustering.
		Only $n=50$ simulations were used for each value of $k$.
	}
\end{figure}

\begin{figure}
	\includegraphics[width=\textwidth]{../_figs/Climatology.pdf}
	\caption{
		NDJF monthly circulation climatology over the study domain and from 1980 to 2010.
    First row shows the \SI{850}{\hecto\pascal} streamfunction $\psi$ climatology, and second row shows rainfall climatology.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/WT_Occurrence_Proportion.pdf}
	\caption{
		Occurrence fraction of NDJF weather types for all years (blue) and 2015-16 (orange).
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/ENSO_EOF_Scatter.pdf}
	\caption{
		Pair plots for three time series: NINO3.4, and EOFs 2 and 3 of the \SI{850}{\hecto\pascal} field over the chosen domain.
    All time series are degraded to monthly time steps.
    Across the diagonal, marginal distributions are shown.
    Straight lines represent linear regression estimates.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/MJO_EOF_Scatter}
	\caption{
		Pair plots for four time series: MJO RMM1 and RMM2, and EOFs 2 and 3 of the \SI{850}{\hecto\pascal} field over the chosen domain.
    All time series are at daily time steps.
    Across the diagonal, marginal distributions are shown.
    Straight lines represent linear regression estimates.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/knn_rainfall_EOF.pdf}
	\caption{
    Expected rainfall over two regions conditional on the (simultaneous) values of EOF2 and EOF3 of the \SI{850}{\hecto\pascal} streamfunction $\psi$ shown in the main text.
    The ``Chaco Jet Extension Region'' is defined as extending from \SIrange{299.25}{302.75}{\degree E} and \SIrange{34.75}{30.75}{\degree S} and the ``Lower Paraguay River Basin'' is defined in the main text.
    Both values represent area-averaged rainfall time series.
    Expected rainfall is calculated by using a nearest-neighbors regression with 100 neighbors.
    Because this methodology performs poorly under extrapolation (where there are few observations), the predicted domain is limited to include only regions with high density of observations.
    }
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../_figs/ENSO_MJO_EOF_PRCP_Time_Series.pdf}
	\caption{
		Time series of S2S predictors and their influence on evolution of circulation and rainfall.
    Top row shows two predictors shown to influence the circulation: monthly NINO3.4 index and daily MJO RMM1 index.
    Blue dots show daily observations of MJO RMM1, while the blue line shows the MJO RMM1 time series with a 7-day rolling mean applied.
    Bottom row shows the evolution of EOFS 2 and 3 of the \SI{850}{\hecto\pascal} field.
    As for the MJO RMM1, dots show the daily fields and lines show the fields with a 7-day running mean applied.
    Bottom rain shows time series of daily rainfall over the ``Lower Paraguay River Basin'' and ``Chaco Jet Extension Region'' defined above.
	}
\end{figure}

>>>>

\begin{figure}
  \includegraphics[width=\textwidth]{../figs/PSI_Var_Explained.pdf}
	\caption{
		Variance explained for each EOF loading of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
    Left figure shows the individual variance explained for each EOF and right figure shows cumulative variance explained.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../figs/WT_Classifiability.pdf}
	\caption{
		Classifiability index (see Methods) calculated for several chosen values of $k$, representing the number of weather types used for clustering.
		Only $n=50$ simulations were used for each value of $k$.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../figs/WT_Occurrence_Proportion.pdf}
	\caption{
		Occurrence fraction of NDJF weather types for all years (blue) and 2015-16 (orange).
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../figs/knn_rainfall_EOF.pdf}
	\caption{
    Expected rainfall over two regions conditional on the (simultaneous) values of EOF2 and EOF3 of the \SI{850}{\hecto\pascal} streamfunction $\psi$ shown in the main text.
    The ``Chaco Jet Extension Region'' is defined as extending from \SIrange{299.25}{302.75}{\degree E} and \SIrange{34.75}{30.75}{\degree S} and the ``Lower Paraguay River Basin'' is defined in the main text.
    Both values represent area-averaged rainfall time series.
    Expected rainfall is calculated by using a nearest-neighbors regression with 100 neighbors.
    Because this methodology performs poorly under extrapolation (where there are few observations), the predicted domain is limited to include only regions with high density of observations.
    }
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{../figs/ENSO_MJO_EOF_PRCP_Time_Series.pdf}
	\caption{
		Time series of S2S predictors and their influence on evolution of circulation and rainfall.
    Top row shows two predictors shown to influence the circulation: monthly NINO3.4 index and daily MJO RMM1 index.
    Blue dots show daily observations of MJO RMM1, while the blue line shows the MJO RMM1 time series with a 7-day rolling mean applied.
    Bottom row shows the evolution of EOFS 2 and 3 of the \SI{850}{\hecto\pascal} field.
    As for the MJO RMM1, dots show the daily fields and lines show the fields with a 7-day running mean applied.
    Bottom rain shows time series of daily rainfall over the ``Lower Paraguay River Basin'' and ``Chaco Jet Extension Region'' defined above.
	}
\end{figure}

### 2017-09-22 writeup/paper.tex
Final weather types calculated w/ 500 sims

Streamflow data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
Locations of streamflow gauges are shown in \cref{fig:study-area}.

@@

The climatological value of a field (rainfall or a circulation variable from reanalysis) is defined as the mean, at each grid cell, during all NDJF months in the period $\qty[1980, 2010]$.
The anomaly, for a particular date, is defined as the departure from this climatology.
Taking climatologies against the NDJF combined climatology, rather than against monthly climatologies, is necessary for weather typing (see below) as discontinuities in the climatological field lead to discontinuities in the anomaly field which leads to inconsistent and results which are difficult to interpret.

\subsection{Weather Typing}

A cluster algorithm is used on daily data to help diagnose the mechanisms associated with the rainfall events of interest in this research.

The clustering was performed on the daily NDJF $\SI{850}{\hecto\pascal}$ streamfunction field ($\psi$), calculated from the meridional and zonal wind fields using spherical harmonics as implemented in the \texttt{windspharm} python package \citep{Dawson:2016ws}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).
To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the anomaly field of $\psi$ (see \cref{sec:climatology-anomalies}) was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for 95\% of the total observed variance (supplemental figure S1).
The \texttt{eofs} python package was used to perform this dimension reduction \citep{Dawson:2016ge}.
No additional time filtering was applied to the data before clustering, thus retaining the annual seasonal cycle, inter-annual, sub-seasonal and synoptic weather time scales for diagnosis.
No latitudinal weighting was necessary as the selected domain is relatively small and it does not extend into high latitudes.
Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance.

Next, the $k$-means algorithm was used to assign a single cluster value to each day on record.
The $k$-means technique is a partitioning method that classifies all days in study into a predefined number of clusters.
It minimizes the intracluster sum of variance, $\mathcal{W}$, for a partition $\mathcal{P}$ defined as the Voronoi diagram generated by the classification process:
\begin{equation}
	\mathcal{W} \qty( \mathcal{P} ) = \sum_{i=1}^{k}\sum_{\vb{x} \in \vb{p}_{i}} \abs{\abs{\vb{x} - \vb{p}_i}}
\end{equation}
where $k$ is the total number of classes in the solution, $\abs{\abs{\cdot}}$ is the Euclidan norm, $\vb{p}_i$ the different clusters in the partition and $\vb{x}$ the maps corresponding to cluster $i$.

The $k$-means algorithm is guaranteed to converge to at least a local minimum; to select the best partition, 250 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv}.
The classifiability index of \citet{Michelangeli1995} was then used to select the optimal partition.
Calculation of the classifiability index for several values of $k$ (supplemental figure S2) suggests that states with $k=4,5,6,7,8$ are all reasonable; we chose the solution $k=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.

@@

\Cref{fig:streamflow} shows the streamflow time series at several gauges on the Paraguay River during NDJF 2015-16 in the context of their seasonality and decadal variability.
Because no stage-discharge curves are available, we present only the river stage.
While this is relevant from the perspective of flood damage, flow rates cannot be estimated without these curves (which are also highly variable in time and consequently must be updated frequently).
In a very flat river basin such as the Lower Paraguay River Basin, such one might expect to expect that discharge scales approximately with $\text{stage}^3$.

The most notable feature of \cref{fig:streamflow}(a) is that during November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on,  and Pilar but not at Bah\'ia Negra.
As discussed in \citet{Bravo:2011et,Barros:2004bn}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
However, the three downstream gauges respond to the rainfall forcing with a slow but steady rise.
Despite several very intense storms (see following sections), the streamflow record at Asunci\'on and Pilar (which are downstream of Concepci\'on) indicates relatively little response to individual storms.

@@

\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 95th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with previous findings \citep{Marengo2004,Salio:2007gd}.
At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$.
Along the cold front associated with this system, a low-level northerly jet advects of heat and moisture to the region, favored by a polar trough extending slightly to the east of Paraguay.
This is in agreement with previous studies which find a close correspondence between the exit regions of the low-level jet and the observed rainfall \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987} as the moisture influx, energy influx, and strong vertical wind shear favor instability and intense convection \citep{Marengo2004,Silva2009}.
Next, low-level convergence and upper-level divergence (not shown) associated with the eastward progression of the frontal system further favor rainfall over the Lower Paraguay River Basin.
As the system progresses, the rainfall anomaly approximately follows the anomaly of the low-level wind.

@@

We next use the weather typing algorithm to understand particular circulations and sequences of circulations associated with intense rainfall in the Lower Paraguay River Basin.

The first step of the weather typing algorithm is to identify leading EOFs of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
The EOF laodings are shown in \cref{fig:eof-loading}.
Of these, EOF 1 explains by far the most variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance (supplemental figure S2).
To a first order of approximation, EOF 1 explains the large-scale anomalies induced by passing baroclinic waves (weaker in the West, likely due to the Andes Mountains).
EOF 2 indicates the strength of the low-level jet circulation, while EOFs 3 and 4 control the exit region of this jet.
Given these features in the EOF loadings, it is no surprise that composites of the composite \SI{850}{\hecto\pascal} circulations and rainfall associated with each weather type (\cref{fig:wt-composite}) reveal circulation patterns are associated with synoptic- and regional-scale circulation patterns.

As predicted, there is a close correspondence between the exit regions of the low-level jet and the observed rainfall.
Weather type 1 represnts a ``Chaco Jet'' event \citep{Salio:2002ev} in which the maximum wind speed penetrates past \SI{25}{\degree S}, leading to intense rainfall NE Argentina and Uruguay.
The other weather type associated with a SALLJ is weather type 4, the ``No-Chaco Jet'' event in which the wind turns to the East before \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
Weather types 2 and 3 resemble the weak and intense phases, respectively, of the South Atlantic Convergence Zone \citep[SACZ;][]{Carvalho2004} while weather types 5 and 6 represent southerly and easterly wind, respectively, over the Lower Paraguay River Basin.
\texttt{ANY THOUGHTS ON WT5,6?}

@@

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{./Streamflow.pdf}
	\caption{
		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
		(a) Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
			The seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in
			the \textbf{R} statistical programming environment \citep{Loader1999}.
		(b) Time series of daily stage measurements from 1929 to 2016 at each station.
	}
  \label{fig:streamflow}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../figs/NDJF201516Anomaly.pdf}
	\caption{
    	Monthly composite anomalies observed during NDJF 2015-16.
      Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
      Bottom row shows monthly rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}
  \label{fig:anomalies}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../figs/PSI_EOF_Loadings.pdf}
	\caption{
    	Leading four EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
	}
  \label{fig:eof-loading}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{../figs/WTComposite.pdf}
	\caption{
		Composite anomalies associated with each weather type (column).
    Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
    Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
	}
	\label{fig:wt-composite}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../figs/RainfallWeatherType}
	\caption{
		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines).
		The weather types corresponding to each day are indicated in both color and adjacent label.
		Dashed lines indicate the climatological 50th, 90th, and 99th percentiles of NDJF area-averaged rain over the Lower Paragauy River basin.
	}
  \label{fig:rain-wt}
\end{figure*}

\begin{figure*}	\noindent\includegraphics[width=6.5in]{../figs/LaggedRain.pdf}
	\caption{
		Circulation anomalies associated with intense rainfall (95th percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin shown in \cref{fig:study-area}).
		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of intense rainfall.
		Top row shows anomalies of streamfunction $\psi$ at \SI{850}{\hecto\pascal}.
    Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
	}
  \label{fig:lagged-rain}
\end{figure*}

>>>>

Streamflow data was collected by the Paraguayan Navy and National Administration of Navigation and Ports of Paraguay, and was processed and distributed by the Paraguayan Directorate of Meteorology and Hydrology (DINAC-DMH).
Locations of streamflow gauges are shown in \cref{fig:study-area}.
Because no stage-discharge curves are available, we present only the river stage values.
While this is relevant from the perspective of flood damage, flow rates cannot be estimated without these curves (which vary in time and consequently must be updated frequently).
In a very flat river basin such as the Lower Paraguay River Basin, such one might expect that discharge scales approximately as $\sim O\qty(\text{stage}^3)$.

@@

We define the climatology of a circulation field $X$ (such as \SI{850}{\hecto\pascal} streamfunction) for each julian day using a moving-window approach.
For each julian day $d = 1, \ldots, 366$, the climatological value of the field $\overline{X}(d)$ is defined as the mean value of all observations $X_t, t=1,\ldots,T$ whose julian day $d_t$ satisfies
\begin{equation}
	d_t \in \qty{\bmod \qty(d_i-30, \ldots, d_i + 30, 366)}
\end{equation}
where $\bmod$ is the modulo operator.
The anomaly at a given time is then the observed field $X_i$ minus the climatological value for the corresponding julian day.
Taking the climatology in this way is advantageous as it retains the seasonal cycle of observations while at the same time avoiding discontinuities in the climatology which is desirable for weather typing (see \cref{sec:methods}\ref{sec:weather-typing}).

\subsection{Weather Typing} \label{sec:weather-typing}

A cluster algorithm is used on daily data to diagnose mechanisms associated with the rainfall events of interest in this research.

The clustering was performed on the daily NDJF $\SI{850}{\hecto\pascal}$ streamfunction field ($\psi$), calculated by integrating the meridional and zonal wind fields using spherical harmonics \citep{Dawson:2016ws}, over the domain spanning \SIrange{15}{30}{\degree S} and \SIrange{65}{45}{\degree W} (\cref{fig:study-area}).

To facilitate clustering (which tends to perform poorly in high-dimensional spaces), the NDJF anomaly field of $\psi_{850}$ (see \cref{sec:climatology-anomalies}) was projected  onto its four leading empirical orthogonal functions (EOFs), accounting for over 95\% of the total observed variance (supplemental figure S1), using the \texttt{eofs} python package \citep{Dawson:2016ge}.
The loading patterns of these EOFs are shown in \cref{fig:eof-loading}.
No meridional weighting was necessary as the selected domain is relatively small and does not extend into high latitudes.
Once the EOFs were calculated, the principal component time series were computed for each day and scaled to unit variance as is typical for cluster analysis.

Next, the $K$-means algorithm was used to assign a single cluster value to each day on record using the 4-dimension principal component time series computed in the previous step.
The $K$-means technique is a partitioning method that classifies all days in the study into a predefined number of clusters.
The algorithm proceeds as follows:
\begin{enumerate}
	\item Randomly choose $K$ cluster centers $\mu_1^{(0)}, \ldots, \mu_K^{(0)}$ (where $0$ refers to the $0$th iteration)
	\item Iterate until convergence, indexing each iteration with $j$:
	\begin{enumerate}
		\item Assign each observation (day) $x_i$ to the nearest cluster center; we use the Euclidean distance:
		\begin{equation}
			m_i^{(j+1)} := \arg\min_{k\in 1,\ldots,K} \abs{\abs{x_i - \mu_k^{(j)}}}
		\end{equation}
		\item Recompute the cluster centers as the mean of all points assigned to that cluster
		\begin{equation}
			\mu_k^{(j+1)} := \frac{1}{\abs{\qty{i \big| m_i^{(j+1)}=k}}} \sum_{i | m_i^{(j+1)}=k} x_i
		\end{equation}
		where $\abs{\cdot}$ denotes vector length.
		\item Stop iteration if the change in centroids $\vec{\mu}^{(j+1)} - \vec{\mu}^{(j)}$ is less than a small but non-zero tolerance parameter $\tau$.
	\end{enumerate}
\end{enumerate}
The cluster centroids (weather types) $\mu_k$ produced by the $K$-means algorithm can then be interpreted as a Voronoi decomposition of the phase space into $K$ regions.
Intuitively, this means that an observation is assigned to the weather type whose centroid $\mu_k$ is closest to it in the phase space given by the leading 4 EOFs of $\psi_{850}.$

The $K$-means algorithm is guaranteed to converge to a local minimum of intercluster variance; to select the best partition, 500 simulations were created using the implementation in Python's \texttt{scikit-learn} package \citep{Pedregosa:2012tv} and then the classifiability index of \citet{Michelangeli1995} was then used to select the optimal partition.
This index can be conceptualized as a stability criterion as it selects the partitioning which is most like the other 249 partitions, as defined by Euclidean distance.
Calculation of the classifiability index for several values of $k$ (supplemental figure S2) suggests that states with $k=4,5,6,7,8$ are all reasonable; we chose the solution $k=6$ because the weather types identified are qualitatively similar to those determined over southeastern South America \citep{Munoz2015,Munoz2016} and have an intuitive physical meaning, discussed further in the following sections.

@@

\Cref{fig:streamflow} shows the streamflow time series at several gauges on the Paraguay River during NDJF 2015-16 in the context of their seasonality and decadal variability.
The most notable feature of \cref{fig:streamflow}(a) is that during November and December 2015, the river rose rapidly at Concepci\'on, Asunci\'on,  and Pilar but not at Bah\'ia Negra.
As discussed in \citet{Bravo:2011et,Barros:2004bn}, the location of the Bah\'ia Negra gauge (see \cref{fig:study-area}) in the Pantanal region means that it responds very slowly to rainfall input.
However, the three downstream gauges respond to the rainfall forcing with a slow but steady rise.
Despite several very intense storms (see following sections), the streamflow record at Asunci\'on and Pilar (which are downstream of Concepci\'on) indicates relatively little response to individual storms.

@@

\Cref{fig:lagged-rain} shows time-lagged anomalies up to and after days exceeding the 95th percentile of area-averaged daily rainfall in the lower Paraguay River basin (see \cref{fig:study-area}), and is consistent with previous findings \citep{Marengo2004,Salio:2007gd}.
At $t=-2$ a mid-latitude baroclinic system approaches the South American continent, intensifying and moving to the East at $t=-1,0,1$.
This system interacts with the sub-tropical low and the Andes Mountains (not shown) to produce an anticyclonic anomaly over Brazil.
Along the cold front associated with this system, a low-level northerly jet advects of heat and moisture to the region.
As the system progresses, the jet transitions from predominantly meridional flow (Chaco Jet; $t=-1$ to predominantly zonal flow below approximately \SI{20}{\degree N}.

The rainfall largely follows the pattern of the low-level jet, in agreement with previous studies  \citep{Saulo:2007km,Salio:2007gd,Marengo2004,Velasco1987}.

@@

We next use the weather typing algorithm outlined in \cref{sec:methods}\ref{sec:weather-typing} to understand particular circulations and sequences of circulations associated with intense rainfall in the Lower Paraguay River Basin.

The first step of the weather typing algorithm is to identify leading EOFs of the \SI{850}{\hecto\pascal} streamfunction $\psi$.
The EOF loadings are shown in \cref{fig:eof-loading}.
Of these, EOF 1 explains by far the most variance ($\approx 67\%$) while EOFs 2, 3, and 4 collectively explain approximately $28\%$ of total variance (supplemental figure S1).
To a first order of approximation, EOF 1 explains the large-scale anomalies induced by passing baroclinic waves; the East-West gradient is likely a result of the influence of the Andes Mountains which are a source of friction for low-level flow.
EOF 2 primarily controls the strength of the low-level jet circulation, while EOFs 3 and 4 control the exit region of this jet.
Given these features in the EOF loadings, it is no surprise that composites of the composite \SI{850}{\hecto\pascal} circulations and rainfall associated with each weather type (\cref{fig:wt-composite}) reveal circulation patterns are associated with synoptic- and regional-scale circulation patterns.

Weather type 1 represents a ``Chaco Jet'' event \citep{Salio:2002ev} in which the maximum wind speed penetrates past \SI{25}{\degree S}, leading to intense rainfall NE Argentina and Uruguay.
The other weather type associated with a SALLJ is weather type 4, the ``No-Chaco Jet'' event in which the wind turns to the East before \SI{25}{\degree S}, leading to heavy rainfall over Eastern Paraguay and SW Brazil.
Weather types 2 and 3 resemble the weak and intense phases, respectively, of the South Atlantic Convergence Zone \citep[SACZ;][]{Carvalho2004} while weather types 5 and 6 represent southerly and easterly wind, respectively, over the Lower Paraguay River Basin.
\texttt{ANY THOUGHTS ON WT5,6?}

@@

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{Streamflow.pdf}
	\caption{
		River stage (height; in \si{\meter}) for the Paraguay River at four gauges along the Paraguay River.
		The station names are shortened versions of those shown in \cref{fig:study-area}.
		(a) Seasonality (orange) and time series of 2015-16 observations (black) at each stream gauge.
			The seasonality was fit using local polynomial regression as implemented in the \texttt{locfit} package in
			the \textbf{R} statistical programming environment \citep{Loader1999}.
		(b) Time series of daily stage measurements from 1929 to 2016 at each station.
	}
  \label{fig:streamflow}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../figs/NDJF201516Anomaly.pdf}
	\caption{
    	Monthly composite anomalies observed during NDJF 2015-16.
      Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
			Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
      Bottom row shows monthly-mean rainfall anomalies, in units of \si{\milli\meter\per\day}.
	}
  \label{fig:anomalies}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../figs/PSI_EOF_Loadings.pdf}
	\caption{
    	Leading four EOF loadings of daily NDJF \SI{850}{\hecto\pascal} streamfunction $\psi$ over chosen domain.
			EOFs are calculated after removing the climatology; see \cref{sec:methods}\ref{sec:climatology-anomalies} for details.
	}
  \label{fig:eof-loading}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in,height=9in,keepaspectratio=true]{../figs/WTComposite.pdf}
	\caption{
		Composite anomalies associated with each weather type (column).
		Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
		Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
    Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
	}
	\label{fig:wt-composite}
\end{figure*}

\begin{figure*}
	\noindent\includegraphics[width=6.5in]{../figs/RainfallWeatherType}
	\caption{
		Time series of area-averaged rainfall in the Lower Paraguay River Basin (\cref{fig:study-area}) for each day of NDJF  2015-16 (lines and points).
		The weather types corresponding to each day are indicated in both color and in an adjacent text label.
		Dashed lines indicate the climatological 50th, 90th, and 99th percentiles of NDJF area-averaged rain over the Lower Paragauy River basin.
	}
  \label{fig:rain-wt}
\end{figure*}

\begin{figure*}	\noindent\includegraphics[width=6.5in]{../figs/LaggedRain.pdf}
	\caption{
		Circulation anomalies associated with intense rainfall (95th percentile exceedance of area-averaged rainfall in the Lower Paraguay River Basin shown in \cref{fig:study-area}).
		Lagged composites are shown, by column, for $t = $ \SIlist{-2;-1;0;1}{\day} relative to the date of intense rainfall.
		Top row shows absolute monthly mean values (color) and monthly-mean anomalies (contour) of streamfunction at \SI{850}{\hecto\pascal} $\psi_{850}$.
		Contour interval for anomalies is \SI{1e6}{\meter\squared\per\second}.
    Bottom row shows precipitation anomalies in units of \si{\milli\meter\per\day}.
	}
  \label{fig:lagged-rain}
\end{figure*}
